---
title: 'InterACT project: examining the inter-rater reliability in screening'
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    toc: false
    reference_docx: rmarkdown-styles-reference.docx
bibliography: references.bib
---

```{r setup, include=FALSE}
# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=500)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation

library(dplyr)
library(tidyr)
library(janitor)
library(ggplot2)
library(irrCAC)
library(ordinal)

# get the data, from an agreement exercise
library(readxl)
data = read_excel('data/InterACTV3IRR-BaselineScreening_V2.xlsx', sheet=2) %>%
  select(-`REDCap ID`, -`CFS score`, -`CriSTAL 1`, -`CriSTAL 2`) %>% # remove a few variables
  clean_names() %>%
  filter(!is.na(auditor)) # remove missing rows

# categorical and continuous variables
cat_vars = names(select(data, -auditor, -case_study, -cfs, -"total_cri_stal", -'spict_score')) # list of all categorical variables 
cont_vars = names(select(data, 'cfs', "total_cri_stal", 'spict_score'))

# loop through categorical varables
results =  NULL
for (v in cat_vars){
columns = select(data, case_study, auditor, v) %>%
  group_by( case_study) %>%
  spread(value=v, key=auditor) %>% # subject = rows, raters=columns
  ungroup() %>%
  select(-case_study)
  # pa = pa.coeff.raw(columns) # percent agreement
  gwet = gwet.ac1.raw(columns)$est
  frame = data.frame(var=v, agreement = gwet$coeff.val, CI = gwet$conf.int, type=1)
  results = bind_rows(results, frame)
  # version with no and unknown combined
  columns = select(data, case_study, auditor, v) %>%
    group_by( case_study) %>%
    mutate_if(is.character, function(x) ifelse(x=='No', 'No/Unknown', x)) %>%
    mutate_if(is.character, function(x) ifelse(x=='Unknown', 'No/Unknown', x)) %>%
    spread(value=v, key=auditor) %>% # subject = rows, raters=columns
  ungroup() %>%
  select(-case_study)
  gwet = gwet.ac1.raw(columns)$est
  frame = data.frame(var=v, agreement = gwet$coeff.val, CI = gwet$conf.int, type=2)
  results = bind_rows(results, frame)
}
results = arrange(results, agreement) %>% # arrange from high to low agreement
  mutate(xaxis = 1:n(),
         xaxis = factor(xaxis, levels=1:n(), labels=var),
         score = as.numeric(var %in% c("cri_stal_risk","spict_status"))+1,
         score = factor(score))
```

There are `r nrow(data)` observations from `r length(unique(data$auditor))` auditors.

## Chance-corrected agreement coefficients for categorical variables

```{r, fig.width=10, fig.height=8}
cat_plot = ggplot(data=results, aes(x=xaxis, y=agreement, col=score, shape=factor(type)))+
  geom_point(size=4)+
  scale_shape_manual('`No` and `Unknown`', values=c(4,0), labels=c('Separate','Combined'))+
  scale_color_manual('CRiSTAL/SPICT', values=c('black','red'), labels=c('No','Yes'))+
  coord_flip()+
  theme_bw()+
  xlab('')+
  ylab('Agreement')+
  theme(
    text = element_text(size=19),
    panel.grid.minor.x = element_blank())
cat_plot
```

The results are ordered from low to high agreement. We use Gwet's agreement statistic [@Wongpakaran2013]. An agreement of one is perfect positive agreement and zero is no agreement. The two key positive/negative classifications of CRiSTAL and SPICT are shown in red.

We run two versions of the agreement statistics. One where the answers "no" and "unknown" are kept as separate answers, and one where they are "combined". This is because "no" and "unknown" have the same implication for the CRiSTAL and SPICT scores.

## Agreement coefficients for continuous variables

```{r}
results_cont = NULL
for (v in cont_vars){
  long = select(data, case_study, auditor, v) 
  formula = paste('as.factor(', v, ')~ -1 + (1|case_study) + (1|auditor)', sep = '')
  model1 <- clmm(formula , link = "probit", threshold = "flexible", data=long)
  var_u = as.numeric(model1$ST$case_study^2) # subjects
  var_v = as.numeric(model1$ST$auditor^2) # experts
  rho = var_u / (var_u + var_v + 1)
frame = data.frame(var=v, rho=rho)
  results_cont = bind_rows(results_cont, frame)
}  
results_cont = arrange(results_cont, rho) %>% # arrange from high to low rho
  mutate(xaxis = 1:n(),
         xaxis = factor(xaxis, levels=1:n(), labels=var))
```

```{r, fig.height=3}
cont_plot = ggplot(data=results_cont, aes(x=xaxis, y=rho))+
  geom_point(size=4)+
  coord_flip()+
  theme_bw()+
  xlab('')+
  ylab('Agreement (rho)')+
  scale_y_continuous(limits=c(0,1))+
  theme(
    panel.grid.minor.x = element_blank())
cont_plot
```

The three continuous variables need a different agreement statistic. We use the ordinal method described in @Nelson15. 

# References
