---
title: InterACT - Multiple imputation model for missing ARP
author: Adrian Barnett
output: pdf_document
geometry: margin=2cm
---

\fontfamily{cmss}
\pagestyle{empty} 
\setlength{\parindent}{5mm}


```{r, echo=FALSE, include=FALSE}
###
### Called from 1_impute_arp.R
###

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(dplyr.summarise.inform=FALSE) # drop annoying groups warning
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation

## Section 0: basics ##
library(dplyr)
library(janitor)
library(tidyr)
library(stringr)
library(flextable) # for nice tables
# graphics
library(ggplot2)
theme_set(theme_bw())

## summarise beta parameters from WinBUGS
all_stats = reshape2::melt(bugs$sims.matrix) %>%
  group_by(Var2) %>%
  summarise(n = n(),
            mean = mean(value),
            median = median(value),
            lower = quantile(value, 0.025),
            upper = quantile(value, 1 - 0.025)) %>%
  ungroup() %>%
  mutate(Var2 = as.character(Var2),
         strip = str_remove_all(Var2, '[^0-9|,]'),
         nicevar = case_when(
           str_detect(Var2, '^tau')  ~ 'Intercept precision',
           str_detect(Var2, '^gamma_c')  ~ 'Random intercept',
           str_detect(Var2, '^alpha') ~ 'Intercept',
           str_detect(Var2, '^beta') ~ 'Slopes')
  ) %>%
  filter(Var2 != 'gamma') %>% # using centred version
  separate(col=strip, into=c('category','predictor'), sep=',', convert=TRUE) %>% # split on comma, can ignore warnings
  rename('Variable' = 'Var2')
```

We used a multinomial model to impute the missing ARP which has three categories of: Yes, valid; Yes, invalid; No. We fitted a multinomial model in a Bayesian paradigm. The equations are:

$$Y_{i} \sim \textrm{Multinomial}(p_{i,j}), \qquad i=1,\ldots,N, \, j=1,2,3,$$
$$p_{i,j} = \pi_{i,j} / \sum_{j=1}^3 \pi_{i,j} ,$$
$$\log(p_{i,j}) = \alpha_j + \gamma_{j,\textrm{team}(i)}  + \sum_{r=1}^4 X_{i,r} \beta_{j,r}$$

Where $\textbf{X}$ is a matrix of patient characteristics that were: age, gender, CRiSTAL score and SPICT score. These variables have no missing data and we hypothesised they would be associated with ARP. 

The model included random intercepts per team:

$$\gamma_{j,\textrm{team}(i)} \sim N(0, \sigma^{-2}_{\gamma,j})$$

These intercepts were important as we expected the rates of completed ARPs to vary by team.

To ensure identifiability we set:
$$\alpha_1 = 0; \, \beta_{1,r} = 0 \, \forall\, r; \, \gamma_{1,t} = 0 \, \forall\, t $$.

We used vague priors for all parameters:
$$\alpha_j \sim N(0, 10^5), \forall\, j$$,
$$\beta_{j,r} \sim N(0, 10^5), \forall\, j,r$$,
$$\sigma^{-2}_{\gamma,j} \sim \textrm{Gamma}(0.1, 0.1), \, j=1,2$$.

We used the WinBUGS software (version 1.4.3). We used `r MCMC` samples after the same sized burn-in. Samples were thinned by `r thin`.

# Number and percent of ARP missing by hospital

```{r}
tab = filter(for_imputation) %>% 
      mutate(missing = as.numeric(is.na(existing_arp)),
             missing = factor(missing, levels=0:1, labels=c('No','Yes'))) %>%
  tabyl(hospital, missing) %>%
  adorn_totals() %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns()

#  adorn_percentages('col')
ftab = flextable(tab) %>% 
    theme_box() %>%
    autofit()
ftab
```

The most amount of missing was at RBWH.

# Table of estimates from the WinBUGS model

The table below shows the estimates from the regression model used to impute ARP. The results are the odds ratio and 95% credible interval.

```{r}
alpha_beta = filter(all_stats, str_detect('beta', string=Variable)) %>%
  mutate(Predictor = case_when(
    predictor == 1 ~ 'Age (+5 years)',
    predictor == 2 ~ 'Gender = Female',
    predictor == 3 ~ 'CRiSTAL',
    predictor == 4 ~ 'SPICT'
  ),
  Category = ifelse(category==2, "Yes, invalid", "No"),
  mean = exp(mean),
  lower = exp(lower),
  upper = exp(upper)) %>%
  select(Predictor, Category, mean, lower, upper)
ftab = flextable(alpha_beta) %>% 
    theme_box() %>%
    autofit() %>%
  colformat_double(digits=2)
ftab
```    

Older age and higher CRiSTAL or SPICT scores were all associated with a reduced odds of no ARP.

# Plot of random intercepts per team

```{r}
to_plot = filter(all_stats, str_detect('^gamma_c', string=Variable)) %>%
  mutate(category_nice = ifelse(category==2, "Yes, invalid", "No"))
rplot = ggplot(to_plot, aes(x=predictor, y=exp(mean), ymin=exp(lower), ymax=exp(upper)))+
  scale_x_continuous(breaks=1:T, labels=NULL)+
  scale_y_log10(breaks=c(0.25,0.5,1,2,6))+
  geom_hline(lty=2, yintercept = 1)+
  geom_point()+
  geom_errorbar(width=0)+
  theme_bw()+
  xlab('Team')+
  ylab('Random intercept (odds)')+
  facet_wrap(~category_nice)+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())
rplot
```    

The plot shows the mean and 95% credible interval per team. There were `r T` teams.