---
title: 'InterACT study: hospital-level summary for teams'
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    toc: false
    reference_docx: rmarkdown-style-teams.docx
---

```{r setup, include=FALSE}
# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
source('99_functions.R')
library(diagram)
library(pander)
library(dplyr)
library(stringr)
library(visdat) # for missing data
library(ggplot2)
library(ggupset)  # to plot combination of symptoms
library(summarytools)
# global options for summary tools
st_options(plain.ascii = FALSE,       # Always use this option in Rmd documents
            style = "rmarkdown",        # This too
            round.digits = 0, 
            headings = FALSE, 
            footnote = NA,             # Avoids footnotes which would clutter the results
            subtitle.emphasis = FALSE  # Improves layout with some rmardown themes
) 

# graphics things:
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7")

# get the data
load('data/FullData.RData') # from 0_read_data.R 
# remove pre-covid data
baseline = filter(baseline, admission_date >= as.Date('2020-05-01'))

## for one hospital - to automate
this_hospital = 'TPCH'
this_team = 'General Medicine'
team_var = 'tpch_team'
names(baseline)[names(baseline) == team_var] = 'team' # rename
baseline = filter(baseline, 
                  hospital == this_hospital,
                  team == this_team)
```

Results for `r this_hospital` hospital, Team = `r this_team`.

### Numbers screened by InterACT and at risk 

```{r, results='asis', fig.width=3.5, fig.height=2.7}
N = nrow(baseline)  # starting number
tab1 = table(baseline$baseline_screening_complete) # table of screening complete
ptab1 = round(prop.table(tab1)*100)
baseline = filter(baseline, baseline_screening_complete=='Yes')
tab2 = table(baseline$at_risk) # table of at risk conditional on screening complete
ptab2 = round(prop.table(tab2)*100)
M <- matrix(nrow = 5, ncol = 5, byrow = TRUE, data = 0)
# names
l1 = paste('Total\n', N)
l2 = paste('Screening\nnot complete\n', tab1[1], ' (', ptab1[1], '%)', sep='')
l3 = paste('Screening\ncomplete\n', tab1[2], ' (', ptab1[2], '%)', sep='')
l4 = paste('Not at risk\n', tab2[1], ' (', ptab2[1], '%)', sep='')
l5 = paste('At risk\n', tab2[2], ' (', ptab2[2], '%)', sep='')
names = c(l1, l2, l3, l4, l5)
# transitions
M[2,1] = "' '"
M[3,1] = "' '"
M[4,3] = "' '"
M[5,3] = "' '"
par(mai=rep(0.01,4))
plotmat(M, name = names, pos=c(1,2,2), curve = 0, box.cex = 1, box.lwd = 2,
        box.size = 0.17, box.prop=0.56, arr.pos = 0.5, box.type = 'rect', arr.type = 'triangle', arr.length = 0.2, arr.width = 0.3, cex=0.25, # cex controls text size
        box.col = c("lightblue","green", "green", "yellow", "yellow"))

### now just at risk patients ###
baseline = filter(baseline, 
                  at_risk=='At risk',
                  !is.na(at_risk))
```

All results from now are for the `r nrow(baseline)` at risk.

### Numbers with a documented care directive prior to this hospital admission. For example: Advance Care Plan, Statement of Choices

```{r, results='asis'}
with(baseline, freq(presence_eol_plan, cumul=FALSE, report.nas = FALSE))
```

#### Weekly numbers with a plan over time (average and 95% confidence interval)

```{r, fig.width=7, fig.height=3}
# make weeks
for_plot = mutate(baseline, 
                  year = as.numeric(format(admission_date, '%Y')),
                  week = as.numeric(format(admission_date, '%U'))/53,
                  weekfrac = year + week) %>%
  group_by(weekfrac) %>%
  summarise(n=n(), yes=sum(presence_eol_plan=='Yes')) %>%
  ungroup() %>%
  mutate(p = yes/n,
         se = sqrt(p*(1-p)/n),
         z = qnorm(0.975),
         lower = p - (z*se), # make confidence interval
         upper = p + (z*se),
         p = p*100, # conver to percent
         lower = lower*100,
         upper = upper*100
         )
# x-axis labels 
x.breaks = paste('2020-', sprintf('%02d', 1:12), '-01', sep='') # make first of month
x.breaks = as.Date(x.breaks)
x.breaks = 2020 + season::yrfraction(x.breaks)
# plot
plot_time = ggplot(data=for_plot, aes(x=weekfrac, y=p, ymin=lower, ymax=upper))+
  geom_point(col='burlywood4', size=4)+
  geom_errorbar(width=0, col='burlywood4', size=1.1)+
  scale_x_continuous(breaks=x.breaks, labels=month.abb)+
  theme_bw()+
  ylab('Percent with a care directive')+
  xlab('Time')
plot_time  
```


# SPICT 

### Ten most common combinations

```{r, fig.width=7, fig.height=4}
# get combinations of yes responses
yes_responses <-  dplyr::select(baseline, participant_id, spict.vars)  %>%
  as_tibble() %>%
  tidyr::gather(key='spict', value='response', -participant_id) %>%
  filter(response=='Yes') %>% # just yes
  select(-response) %>%
  mutate(spict = nice.rename.spict(spict)  )
# now make list from group responses
list_resp = group_by(yes_responses, participant_id) %>%
  summarise(spicts = list(spict))
# add zeros
all_ids = unique(baseline$participant_id)
missing.ids = all_ids[all_ids %in% list_resp$participant_id == FALSE]
zeros = as_tibble(data.frame(participant_id=missing.ids))
list_resp = bind_rows(list_resp, zeros)
# plot
cplot = ggplot(list_resp, aes(x = spicts)) +
    geom_bar(aes(y=..count../sum(..count..)), fill = "indianred3") +
    theme_bw() +
    xlab("") +
    ylab("Proportion of patients") +
    scale_x_upset(n_intersections = 10, n_sets=10)+ # top ten
    g.theme+
  theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 90, unit = "pt"))
cplot
```

### Numbers positive

The threshold for SPICT positive is `r thresholds$spict_threshold` or more.

```{r, results='asis'}
baseline = mutate(baseline, 
                  `SPICT positive` = as.numeric(spict_score >= thresholds$spict_threshold),
                  `SPICT positive` = factor(`SPICT positive`, levels=0:1, labels=c('No','Yes')))
with(baseline, freq(`SPICT positive`, cumul = FALSE, report.nas = FALSE))
```

# CriSTAL 

### Ten most common combinations

```{r, fig.width=7, fig.height=5}
# get combinations of yes responses
yes_responses <-  dplyr::select(baseline, participant_id, cristal.vars)  %>%
  mutate(
    #age = ifelse(age>=75, 'Yes', 'No'), # convert to binary - no longer using age in cristal
         cristal_cfs_score1 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold, 'Yes', 'No'), # frailty score over 5
         cristal_cfs_score2 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold2, 'Yes', 'No')) %>%  # frailty score over 7
  select(-cristal_cfs_score) %>% # no longer needed
  mutate_if(is.factor, as.character) %>% # convert all yes/no factors to characters
  as_tibble() %>%
  tidyr::gather(key='cristal', value='response', -participant_id) %>%
  filter(response=='Yes') %>% # just yes
  select(-response) %>%
  mutate(cristal = nice.rename.cristal(cristal)  )
# now make list from group responses
list_resp = group_by(yes_responses, participant_id) %>%
  summarise(cristals = list(cristal))
# add zeros
all_ids = unique(baseline$participant_id)
missing.ids = all_ids[all_ids %in% list_resp$participant_id == FALSE]
zeros = as_tibble(data.frame(participant_id=missing.ids))
list_resp = bind_rows(list_resp, zeros)
# plot
# to do, colour bar by threshold?
cplot = ggplot(list_resp, aes(x = cristals)) +
    geom_bar(aes(y=..count../sum(..count..)), fill = "#6855CD") +
    theme_bw() +
    xlab("") +
    ylab("Proportion of patients") +
    scale_x_upset(n_intersections = 10, n_sets=10) # top ten
cplot
```

### Numbers positive

A CriSTAL positive score is `r thresholds$cristal_threshold` or above.

```{r, results='asis'}
baseline = mutate(baseline, 
                  `CriSTAL positive` = as.numeric(cristal_score >= thresholds$cristal_threshold),
                  `CriSTAL positive` = factor(`CriSTAL positive`, levels=0:1, labels=c('No','Yes')))
with(baseline, freq(`CriSTAL positive`, cumul=FALSE, report.nas = FALSE))
```
