---
title: 'InterACT project: overall statistical analyses for all three hospitals. Version not split by hospital.'
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    toc: true
    toc_depth: 2
    reference_docx: rmarkdown-styles-reference.docx
---

```{r setup, include=FALSE}
# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
source('99_functions.R')
library(cmprsk) # for cumulative incidence
library(broom) # for regression models
library(dplyr)
library(tidyr)
library(stringr)
library(visdat) # for missing data
library(flextable)
library(summarytools)
# global options for summary tools
st_options(plain.ascii = FALSE,       # Always use this option in Rmd documents
            style = "rmarkdown",        # This too
            round.digits = 0, 
            headings = FALSE, 
            footnote = NA,             # Avoids footnotes which would clutter the results
            subtitle.emphasis = FALSE  # Improves layout with some rmarkdown themes
) 

## get the data
data_to_use = 'real'
file_plus = '' # for figures
if(data_to_use=='real'){load('data/FullData.RData')} # from 0_read_data.R 
if(data_to_use=='dummy'){load('data/DummyData.RData'); file_plus='_dummy'} # dummy data from 0_make_dummy_data.R 
# just use version three
baseline = filter(baseline, redcap_version==3)
complete = filter(complete, redcap_version==3)
care_directive = filter(care_directive, redcap_version==3)
palliative_care_referral = filter(palliative_care_referral, redcap_version==3)
clinicianled_review = filter(clinicianled_review, redcap_version==3)

# graphics things:
library(ggplot2)
library(ggalluvial) # for Sankey diagram
library(ggcorrplot) # for correlation matrix
library(ggupset)  # to plot combination of symptoms
library(survminer) # for survival plots
library(gridExtra)
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7")
```


## Overall numbers

##### All baseline screening completed  

```{r, results='asis'}
with(baseline, freq(baseline_screening_complete, round.digits=0, cumul = FALSE, report.nas =FALSE))
```

##### Hospitals, at risk patients for those with baseline screening complete

```{r, results='asis'}
baseline = filter(baseline, baseline_screening_complete=='Yes')
with(baseline, freq(at_risk, round.digits=0, cumul = FALSE, report.nas =FALSE))
```

From now on, all results are for those with baseline screening completed

The table shows the numbers by hospital that were either CriSTAL or SPICT positive.
The total column shows the total number of patients with complete screening.

# Patients screened 


### Overall numbers by study time period

The table shows the numbers in each hospital by study time period.

```{r}
tab = group_by(baseline, int_time) %>%
  tally() %>%
  rename(`Time` = 'int_time')
ftab = flextable(tab) %>%
  theme_box() %>%
  merge_v(j=1) %>%
  autofit()
ftab
```

### At-risk numbers by study time period

The table shows the numbers in each hospital by study time period and the number (percent) at risk.

```{r}
tab = group_by(baseline, int_time, at_risk) %>%
  tally() %>%
  group_by(int_time) %>%
  mutate(percent = prop.table(n)*100,
         cell = paste(n, ' (', round(percent), ')', sep='')) %>%
  select(-n, -percent) %>%
  spread(at_risk, cell) %>%
  rename(`Time` = 'int_time')
ftab = flextable(tab) %>%
  theme_box() %>%
  autofit()
ftab
```

### Admission time

```{r, out.width='90%'}
hplot = ggplot(data=baseline, aes(x=admission_time))+
  geom_histogram(alpha=0.7, breaks=seq(0,24,1), fill='darkseagreen1', colour='grey77')+
  xlab('Time of day')+
  ylab("Count")+
  scale_x_continuous(breaks=seq(0,24,6))+
  g.theme
hplot
filename = paste('figures/all_sites_admission_time', file_plus, '.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(hplot)
invisible(dev.off())
```

The majority of patients were admitted in the afternoon and early evening.

### Times between admission and at-risk screening

We examine the time between the patients' admission date/time to the current ward and the InterACT screening. We use two date/times for screening:

* The earliest date/time from the four screening forms: demographics, clinical criteria, functional status and comorbidities. This is an estimate of when screening started.
* The date/time the screening completion form was completed. This is an estimate of when screening ended.

##### Box plots

```{r, fig.width=8}
# add screening form completion date
for_merge = select(complete, participant_id, start_date_time_dcharg) # just date
to_table = left_join(baseline, for_merge, by='participant_id') %>%
  mutate(
    diff_start = (as.numeric(median_form) - as.numeric(admission_datetime))/(60*60*24), # in days from earliest date
    diff_end = (as.numeric(start_date_time_dcharg) - as.numeric(admission_datetime))/(60*60*24), # in days from latest date
    admission_datetime = as.POSIXct(as.numeric(admission_datetime), tz='Australia/Brisbane', origin='1970-01-01'),
    start_date_time_dcharg = as.POSIXct(as.numeric(start_date_time_dcharg), tz='Australia/Brisbane', origin='1970-01-01')
         ) %>% 
  select(participant_id, hospital, admission_datetime, start_date_time_dcharg, diff_start, diff_end, at_risk) %>%
  group_by(participant_id, at_risk) %>%
  gather(`diff_start`, `diff_end`, key='difference', value='var') %>%
  ungroup() %>%
  mutate(difference = ifelse(str_detect(string=difference, pattern='start'), 'Start', 'End'))
# (checks)
check = filter(to_table, var>60) %>%
  select(-at_risk, -var)
write.csv(check, file='checks/admission_dates.csv', row.names = FALSE, quote=FALSE)
# plot
hplot = ggplot(data=to_table, aes(x=var, fill=at_risk))+
  geom_boxplot(width=0.1)+
  scale_fill_manual(NULL, values=cbPalette)+
  xlab('Difference from ward admissision to screening complete (days)')+
  ylab('')+
  facet_wrap(~difference)+
  scale_y_continuous(breaks=NULL)+
  scale_x_log10()+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
hplot
```

The time axis is on a log scale (base 10) because of the strong positive skew in the difference in times.

##### Summary statistics

```{r}
stats = group_by(to_table, at_risk, difference) %>%
  summarise(n = n(),
            missing = sum(is.na(var)),
            q1 = roundz(quantile(var, prob=0.25, na.rm=TRUE),1),
            median = roundz(median(var, na.rm=TRUE),1),
            q3 = roundz(quantile(var, prob=0.75, na.rm=TRUE),1)) %>%
  rename('At risk' = 'at_risk',
         'Screening time' = 'difference')
ft = flextable(stats)
autofit(ft)
```

The statistics are the number observed, number missing, median and inter-quartile range.

# Demographics

### Patient age

The plot is split by CriSTAL/SPICT positive.

```{r, out.width='100%', fig.width=7}
p <- ggplot(filter(baseline, !is.na(at_risk)), aes(x=factor(at_risk), y=age, fill=factor(at_risk))) + 
  scale_fill_manual('', values=cbPalette)+
  geom_violin(alpha=0.6, bw=2)+
  geom_boxplot(width=0.1)+
  xlab('')+
  ylab('Age')+
  g.theme+
  theme(legend.position = 'none')
p
filename = paste('figures/all_sites_age', file_plus, '.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(p)
invisible(dev.off())
```

### Patient gender

```{r, results='asis'}
freq(baseline$pt_sex, cumul=FALSE, report.nas = FALSE)
```

# SPICT 

The SPICT score has a range from 0 to 6 and uses the following six variables:

* Unplanned hospital admission
* Performance status is poor or deteriorating
* Depends on others for care
* Progressive weight loss; remains under weight; low muscle mass
* Persistent symptoms despite optimal treatment of underlying conditions
* Patient (or family) asks for palliative care

Each of the six variables contributes equally to the score. Unknown or missing values are assumed to be "No".

### Correlation matrix of SPICT variables

```{r spict.corr}
# function to make yes/no into number
z_makenum <- function(observed) {
  result = ifelse(observed=='Yes', 1, 0)
  return(result)
}
# estimate correlation matrix
to.corr = select(baseline, spict.vars) %>% # these are just yes/no/unknown
  mutate_all(funs(z = z_makenum(.))) %>%
  select(ends_with('_z')) %>%
  rename_at(vars(ends_with("_z")),funs(str_replace(.,"_z","")))  # remove _z from name
corr <- cor(to.corr, use = 'pairwise')
colnames(corr) = row.names(corr) = nice.rename.spict(colnames(corr)) # rename
# plot
diag(corr) = NA # blank diagonals
ggcorrplot(corr) 
# for text around strongest correlation
max.corr = max(corr[lower.tri(corr, diag=FALSE)])
vars = which(corr==max.corr, arr.ind = T)
```

The plot shows the observed correlation between SPICT variables. Correlations for two variables range from -1 if they are perfectly negatively associated, to 0 if there is no association, to +1 if they are perfectly positively associated.

The strongest positive correlation is `r round(max.corr,2)` between "`r row.names(vars)[1]`" and "`r row.names(vars)[2]`". There were no negative correlations.

### Tables of SPICT variables

```{r, results='asis'}
## big summary table
# create summary stats for table / plot
spict = select(baseline, "participant_id", all_of(spict.vars)) %>%
  tidyr::gather(key='SPICT', value='Response', -`participant_id`) %>%
  mutate(SPICT = nice.rename.spict(SPICT)) # neaten variable
with(spict, ctable(SPICT, Response, dnn=c('SPICT',''), round.digits=0, useNA='no'))
```


### SPICT score

```{r, fig.width=8, results='asis'}
to.bar = group_by(baseline, spict_score) %>%
  summarise(count = n()) %>%
  mutate(fill = as.factor(spict_score >= thresholds$spict_threshold))
spict.bar = ggplot(data=to.bar, aes(x=spict_score, y=count, fill=fill))+ # colour bars by positive
  xlab('SPICT score')+
  ylab('Count')+
  scale_x_continuous(breaks=0:6)+
  geom_bar(stat='identity')+
  scale_fill_manual(NULL, values=c('skyblue','dark blue'))+
  g.theme+
  theme(legend.position = 'none')
spict.bar
filename = paste('figures/all_sites_spict_bar', file_plus, '.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(spict.bar)
invisible(dev.off())
```

The darker blue bars show the SPICT-positive patients.

## SPICT positive

A patient is SPICT positive if their score is `r thresholds$spict_threshold` or greater.

```{r, results='asis'}
baseline = mutate(baseline, 
                  `SPICT positive` = as.numeric(spict_score >= thresholds$spict_threshold),
                  `SPICT positive` = factor(`SPICT positive`, levels=0:1, labels=c('No','Yes')))
with(baseline, freq(`SPICT positive`, round.digits=0, cumul = FALSE, report.nas =FALSE))
```

## Common combinations for SPICT

Here we show the most common combinations of symptoms for patients.

```{r, fig.width=7}
# get combinations of yes responses
yes_responses <-  dplyr::select(baseline, participant_id, spict.vars)  %>%
  as_tibble() %>%
  tidyr::gather(key='spict', value='response', -participant_id) %>%
  filter(response=='Yes') %>% # just yes
  select(-response) %>%
  mutate(spict = nice.rename.spict(spict)  )
# now make list from group responses
list_resp = group_by(yes_responses, participant_id) %>%
  summarise(spicts = list(spict))
# add zeros
all_ids = unique(baseline$participant_id)
missing.ids = all_ids[all_ids %in% list_resp$participant_id == FALSE]
zeros = as_tibble(data.frame(participant_id=missing.ids))
list_resp = bind_rows(list_resp, zeros)
# plot
cplot = ggplot(list_resp, aes(x = spicts)) +
    geom_bar(aes(y=..count../sum(..count..)), fill = "indianred3") +
    g.theme+
    xlab("SPICT combinations") +
    ylab("Proportion of patients") +
    scale_x_upset(n_intersections = 20) # top twenty
cplot
# export to figure
filename = paste('figures/all_sites_spict_upset', file_plus, '.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(cplot)
invisible(dev.off())
```

# CriSTAL 

The CriSTAL score has a range in this study from 1 to 18. It uses the following 16 variables:

* Age over 75 (_all patients in the InterACT study_)
* Admitted via Emergency Department
* Admitted From nursing home or supported accommodation
* Previous hospitalisation in last 12 months
* ICU admission
* Clinical frailty score (two thresholds at 5+ and 7+)
* Advanced cancer
* Proteinuria
* Chronic kidney disease
* Abnormal ECG
* Acute myocardial infarction
* Chronic heart failure
* COPD
* Stroke
* Cognitive impairment
* Liver disease

And 8 clinical criteria:

* Decreased level of consciousness (LOC) Glasgow Coma Scale (GCS) decreased more than 2 compared with baseline for this person
* Systolic Blood Pressure less than 90 mmHg
* Respiratory Rate less than 5 or greater than 30 bpm	
* Pulse Rate less than 40 or greater than 140 bpm	
* Oxygen required or oxygen saturation less than 90%	
* Hypoglycaemia (BGL 1.0 to 4.0 mmol/L)
* Repeat or prolonged seizures (any type) More than one in 24hrs (including prior to hospital admission) or duration greater than 5 min	
* Low urine output (less than 15 ml/hr or less than 0.5 ml/kg/hr)	

We use two thresholds for the clinical frailty score at 5 or more and 7 or more.

### Correlation matrix of CriSTAL variables

```{r cristal.corr, fig.width=7, fig.height=7}
# function to make yes/no into number
z_makenum <- function(observed) {
  result = ifelse(observed=='Yes', 1, 0)
  return(result)
}
# estimate correlation matrix
to.corr = select(baseline, cristal.vars) %>%
  mutate_at(vars(!ends_with("_cfs_score")), funs(z = z_makenum(.))) %>% # only mutate binary variables
  select(ends_with('_z'), 'cristal_cfs_score') %>%
  rename_at(vars(ends_with("_z")), funs(str_replace(.,"_z",""))) # remove _z from name
corr <- cor(to.corr, use = 'pairwise')
colnames(corr) = row.names(corr) = nice.rename.cristal(colnames(corr)) # rename
colnames(corr)[colnames(corr) == "Frailty score >= 6"] = 'Frailty score'
row.names(corr)[row.names(corr) == "Frailty score >= 6"] = 'Frailty score'
# plot
diag(corr) = NA # blank diagonals
ggcorrplot(corr, tl.cex=8) # cex to reduce text size
# for text around strongest correlations
max.corr = max(corr[lower.tri(corr, diag=FALSE)])
vars.pos = which(corr==max.corr, arr.ind = T)
min.corr = min(corr[lower.tri(corr, diag=FALSE)])
vars.neg = which(corr==min.corr, arr.ind = T)
```

The plot shows the observed correlation between CRiSTAL variables. The strongest positive correlation is `r round(max.corr,2)` between "`r row.names(vars.pos)[1]`" and "`r row.names(vars.pos)[2]`". The strongest negative correlation is `r round(min.corr,2)` between "`r row.names(vars.neg)[1]`" and "`r row.names(vars.neg)[2]`". 

The clinical frailty score is continuous, whereas the other variables are binary. To make the binary variables numeric, we group 'no' and 'unknown' together as zero, and compare them with 'yes' as one.

We did not include age in the correlation plot as every patient is over 75.

### Tables of CriSTAL variables

```{r, results='asis'}
## big summary table
# create summary stats for table / plot
cristal = select(baseline, "participant_id", 'hospital', all_of(cristal.vars)) %>%
  mutate(
    #age = ifelse(age>=75, 'Yes', 'No'), # convert to binary , no longer using age in cristal
         cristal_cfs_score1 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold, 'Yes', 'No'), # frailty score over 5,
         cristal_cfs_score2 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold2, 'Yes', 'No') # frailty score over 7,
         ) %>% 
  select(-cristal_cfs_score) %>% # no longer need score
  mutate_if(is.factor, as.character) %>% # convert all yes/no factors to characters
  tidyr::gather(key='CRISTAL', value='Response', -`participant_id`, -`hospital`) %>%
  mutate(CRISTAL = nice.rename.cristal(CRISTAL)) # neaten variable
# 
with(cristal, ctable(CRISTAL, Response, dnn=c('CriSTAL Variable',''), round.digits=0, useNA='no'))
```

### CriSTAL score

```{r, fig.width=8}
#
MaxC = max(baseline$cristal_score) # for x-axis limit
#
to.bar = group_by(baseline, cristal_score) %>%
  summarise(count = n()) %>%
  mutate(fill = as.factor(cristal_score >= thresholds$cristal_threshold))
cristal.bar = ggplot(data=to.bar, aes(x=cristal_score, y=count, fill=fill))+ # colour bars by positive
  xlab('CriSTAL score')+
  ylab('Count')+
  scale_x_continuous(breaks=0:MaxC)+
  geom_bar(stat='identity')+
  scale_fill_manual(NULL, values=c('skyblue','dark blue'))+
  g.theme+
  theme(legend.position = 'none')
filename = paste('figures/all_sites_cristal_bar', file_plus, '.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(cristal.bar)
invisible(dev.off())
```

The darker blue bars show the CriSTAL-positive patients with a score of `r thresholds$cristal_threshold` or greater. 


### CriSTAL score summary statistics

```{r, cristal.summary, results='asis'}
descr(baseline$cristal_score, stats = c('N.valid','mean','sd','min','med','max'), round.digits =1, transpose = TRUE)
```

### CriSTAL positive

```{r, results='asis'}
baseline = mutate(baseline, 
                  `CriSTAL positive` = as.numeric(cristal_score >= thresholds$cristal_threshold),
                  `CriSTAL positive` = factor(`CriSTAL positive`, levels=0:1, labels=c('No','Yes')))
with(baseline, freq(`CriSTAL positive`, round.digits=0, cumul = FALSE, report.nas =FALSE))
```

A CriSTAL positive score is `r thresholds$cristal_threshold` or above, and we apply this for the entire cohort.

## Common combinations for CriSTAL 

Here we show the twenty most common combinations of sign and symptoms for screened patients. 
We do not show age over 75 in this plot, because everyone in our sample is over this age.
If a CriSTAL variable is not shown, it means it was not used in any of the top twenty combinations.

```{r, fig.width=8}
# get combinations of yes responses
yes_responses <-  dplyr::select(baseline, participant_id, cristal.vars)  %>%
  mutate(
    #age = ifelse(age>=75, 'Yes', 'No'), # convert to binary - no longer using age in cristal
         cristal_cfs_score1 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold, 'Yes', 'No'), # frailty score over 5
         cristal_cfs_score2 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold2, 'Yes', 'No')) %>%  # frailty score over 7
  select(-cristal_cfs_score) %>% # no longer needed
  mutate_if(is.factor, as.character) %>% # convert all yes/no factors to characters
  as_tibble() %>%
  tidyr::gather(key='cristal', value='response', -participant_id) %>%
  filter(response=='Yes') %>% # just yes
  select(-response) %>%
  mutate(cristal = nice.rename.cristal(cristal)  )
# now make list from group responses
list_resp = group_by(yes_responses, participant_id) %>%
  summarise(cristals = list(cristal))
# add zeros
all_ids = unique(baseline$participant_id)
missing.ids = all_ids[all_ids %in% list_resp$participant_id == FALSE]
zeros = as_tibble(data.frame(participant_id=missing.ids))
list_resp = bind_rows(list_resp, zeros)
# plot
# to do, colour bar by threshold?
cplot = ggplot(list_resp, aes(x = cristals)) +
    geom_bar(aes(y=..count../sum(..count..)), fill = "#6855CD") +
    g.theme+
    xlab("CriSTAL combinations") +
    ylab("Proportion of patients") +
    scale_x_upset(n_intersections = 20) # top twenty
cplot
# export to figure
filename = paste('figures/all_sites_cristal_upset', file_plus, '.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(cplot)
invisible(dev.off())
```

### Frailty score

The plot is split by CriSTAL/SPICT positive.

```{r, out.width='100%', fig.width=7}
p <- ggplot(filter(baseline, !is.na(at_risk)), aes(x=factor(at_risk), y=cristal_cfs_score, fill=factor(at_risk))) + 
  scale_fill_manual('', values=cbPalette)+
  geom_violin(alpha=0.6, bw=2)+
  geom_boxplot(width=0.1)+
  xlab('')+
  ylab('Clinical frailty score')+
  g.theme+
  theme(legend.position = 'none')
p
filename = paste('figures/all_sites_cfs', file_plus, '.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(p)
invisible(dev.off())
```
