---
title: 'InterACT project: outcomes 4, 5 and 6 by team using survival plots'
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    toc: false
    reference_docx: rmarkdown-styles-reference.docx
bibliography: references.bib
---

```{r setup, include=FALSE}
# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400, dev = "ragg_png")
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
source('99_functions.R')
library(cmprsk) # for cumulative incidence
library(dplyr)
library(tidyr)
library(janitor) # for tables with column totals
library(stringr)
library(flextable)

# graphics things:
library(ragg) # because of ribbon issue
library(ggplot2)
library(survminer) # for survival plots
library(gridExtra)
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7") # colour-blind palette
pred_time = 21 # for right limit of plots

# start with the unscrambled data
source('1_not_scramble_data.R')

# scramble intervention (turn off -- FALSE -- when ready)
scramble = FALSE # 
if(scramble == TRUE){
  source('1_scramble_data.R')
}

# create small subset of variables to add to outcome data
data_to_add = select(baseline, participant_id, pt_sex, age, team)
```

This report examines outcomes 4, 5 and 6 by team.


# Clinician-led review discussion

Here we use time-to-event analysis to examine the time to the first clinical-led discussion The plot below shows the cumulative probability for the first clinical-led discussion by intervention phase.

```{r, include=FALSE}
## prepare data for cumulative and instantaneous survival models
# exclude prior:
for_model_care_review = filter(survival_data,
                  outcome == 'care_review',
                  int_time != 'Establishment', # excluded from analysis
                  event != 'Prior') %>%
  mutate(  event = as.character(event), # needed for ifelse below
           event = ifelse(event== 'No', 'Censored', event)) # change "No" to "censored" for this analysis. Competing risks are death/event

# add variables from baseline data that are not in outcome data (age, sex)
for_model_care_review = left_join(for_model_care_review, data_to_add, by='participant_id')

## numbers excluded
# a) establishment
n_excluded_establishment = nrow(filter(survival_data, 
                               outcome == 'care_review',
                               int_time == 'Establishment'))
# b) prior
n_excluded_prior = nrow(filter(survival_data, 
                               outcome == 'care_review',
                               int_time != 'Establishment', # avoid double-counting of those excluded for this reason
                               event == 'Prior'))
```

##### Cumulative probability plot (clinical-led discussion)

```{r, include=FALSE}
# cumulative probability plot
# plot in each hospital - change after scrambling
plot_data = NULL
for (hosp in c('RBWH','TPCH','GCUH')){
  teams = filter(for_model_care_review, hospital==hosp) %>% # get all teams
    pull(team) %>%
    unique()
for (this_team in teams){
  this_data = filter(for_model_care_review, hospital ==hosp, team==this_team) %>%
    mutate(eventSimple = ifelse(event!='Yes', 'Censored', 'Yes')) # everything not yes is censored
  cum_inc = with(this_data, cuminc(ftime=time, fstatus=eventSimple, group=int_time, cencode='Censored'))
  g = ggcompetingrisks(cum_inc,
                       xlab = 'Days',
                       ylab = 'Cumulative probability',
                       multiple_panels = FALSE,
                       title = '',
                       cumcensor = TRUE, # does not work
                       legend.title='',
                       risk.table = TRUE, # does not work
                       conf_int = TRUE) # does not work
  # combine data
  this_plot = mutate(g$data, hospital=hosp, team=this_team)
  plot_data = bind_rows(plot_data, this_plot)
} # end of team loop
} # end of hospital loop
# add CI
to_plot = mutate(plot_data,
                 z = qnorm(0.975),
                 lower  = est - (z*sqrt(var)),
                 upper  = est + (z*sqrt(var)),
                 group = ifelse(group=='Usual', 'Usual care', group))
```

### RBWH

```{r, fig.width=8, fig.height=7}
rbwh = filter(to_plot, hospital=='RBWH')
# redo plot for greater control
alt_plot = ggplot(rbwh, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0,1)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.85, 0.15)) +
  facet_wrap(~team)
alt_plot
```

### TPCH

```{r, fig.width=8, fig.height=6}
tpch = filter(to_plot, hospital=='TPCH')
# redo plot for greater control
alt_plot = ggplot(tpch, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0,1)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.85, 0.15)) +
  facet_wrap(~team)
alt_plot
```

### GCUH

```{r, fig.width=8, fig.height=7}
gcuh = filter(to_plot, hospital=='GCUH')
# redo plot for greater control
alt_plot = ggplot(gcuh, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0, 0.8)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.9, 0.12)) +
  facet_wrap(~team)
alt_plot
```

###### page break

# Review of care directives documents

The results in this section only use the first care directive per patient. For patients without a review, we used their final completed form.

```{r, include=FALSE}
## prepare data for cumulative and instantaneous survival models
# exclude prior:
for_model_care_directive = filter(survival_data,
                  outcome == 'care_directive',
                  int_time != 'Establishment', # excluded from analysis
                  event != 'Prior') %>%
  mutate(  event = as.character(event), # needed for ifelse below
           event = ifelse(event== 'No', 'Censored', event)) # change "No" to "censored" for this analysis. Competing risks are death/event

# add variables from baseline data
for_model_care_directive = left_join(for_model_care_directive, data_to_add, by='participant_id')
```

```{r, include=FALSE}
## cumulative probability plot
# plot in each hospital - change after scrambling
plot_data = NULL
for (hosp in c('RBWH','TPCH','GCUH')){
  teams = filter(for_model_care_directive, hospital==hosp) %>% # get all teams
    pull(team) %>%
    unique()
for (this_team in teams){
  this_data = filter(for_model_care_directive, hospital ==hosp, team==this_team) %>%
    mutate(eventSimple = ifelse(event!='Yes', 'Censored', 'Yes')) # everything not yes is censored
  cum_inc = with(this_data, cuminc(ftime=time, fstatus=eventSimple, group=int_time, cencode='Censored'))
  g = ggcompetingrisks(cum_inc,
                       xlab = 'Days',
                       ylab = 'Cumulative probability',
                       multiple_panels = FALSE,
                       title = '',
                       cumcensor = TRUE, # does not work
                       legend.title='',
                       risk.table = TRUE, # does not work
                       conf_int = TRUE) # does not work
  # combine data
  this_plot = mutate(g$data, hospital=hosp, team=this_team)
  plot_data = bind_rows(plot_data, this_plot)
}
}
# add CI
to_plot = mutate(plot_data,
                 z = qnorm(0.975),
                 lower  = est - (z*sqrt(var)),
                 upper  = est + (z*sqrt(var)),
                 group = ifelse(group=='Usual', 'Usual care', group))
```

### RBWH

```{r, fig.width=8, fig.height=7}
rbwh = filter(to_plot, hospital=='RBWH')
# redo plot for greater control
alt_plot = ggplot(rbwh, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0,1)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.85, 0.15)) +
  facet_wrap(~team)
alt_plot
```

### TPCH

```{r, fig.width=8, fig.height=6}
tpch = filter(to_plot, hospital=='TPCH')
# redo plot for greater control
alt_plot = ggplot(tpch, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0, 0.75)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.9, 0.12)) +
  facet_wrap(~team)
alt_plot
```

### GCUH

```{r, fig.width=8, fig.height=7}
gcuh = filter(to_plot, hospital=='GCUH')
# redo plot for greater control
alt_plot = ggplot(gcuh, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0, 0.8)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.9, 0.1)) +
  facet_wrap(~team)
alt_plot
```

###### page break

# Palliative care referral

The results in this section only use the first palliative care referral per patient, or for patients without any referrals in their completed forms.

```{r, include=FALSE}
## prepare data for cumulative and instantaneous survival models
# exclude prior:
for_model_palliative = filter(survival_data,
                  outcome == 'palliative_care_referral',
                  int_time != 'Establishment', # excluded from analysis
                  event != 'Prior') %>%
  mutate(  event = as.character(event), # needed for ifelse below
           event = ifelse(event== 'No', 'Censored', event)) # change "No" to "censored" for this analysis. Competing risks are death/event

# add variables from baseline data (age, sex)
for_model_palliative = left_join(for_model_palliative, data_to_add, by='participant_id')
```

```{r, include=FALSE}
## cumulative probability plot
# plot in each hospital - change after scrambling
plot_data = NULL
for (hosp in c('RBWH','TPCH','GCUH')){
  teams = filter(for_model_palliative, hospital==hosp) %>% # get all teams
    pull(team) %>%
    unique()
for (this_team in teams){
  this_data = filter(for_model_palliative, hospital ==hosp, team==this_team) %>%
    mutate(eventSimple = ifelse(event!='Yes', 'Censored', 'Yes')) # everything not yes is censored
  cum_inc = with(this_data, cuminc(ftime=time, fstatus=eventSimple, group=int_time, cencode='Censored'))
  g = ggcompetingrisks(cum_inc,
                       xlab = 'Days',
                       ylab = 'Cumulative probability',
                       multiple_panels = FALSE,
                       title = '',
                       cumcensor = TRUE, # does not work
                       legend.title='',
                       risk.table = TRUE, # does not work
                       conf_int = TRUE) # does not work
  # combine data
  this_plot = mutate(g$data, hospital=hosp, team=this_team)
  plot_data = bind_rows(plot_data, this_plot)
}
}
# add CI
to_plot = mutate(plot_data,
                 z = qnorm(0.975),
                 lower  = est - (z*sqrt(var)),
                 upper  = est + (z*sqrt(var)),
                 group = ifelse(group=='Usual', 'Usual care', group))
```

### RBWH

```{r, fig.width=8, fig.height=7}
rbwh = filter(to_plot, hospital=='RBWH')
# redo plot for greater control
alt_plot = ggplot(rbwh, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0,0.6)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.85, 0.15)) +
  facet_wrap(~team)
alt_plot
```

### TPCH

```{r, fig.width=8, fig.height=6}
tpch = filter(to_plot, hospital=='TPCH')
# redo plot for greater control
alt_plot = ggplot(tpch, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0,0.3)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.15, 0.9)) +
  facet_wrap(~team)
alt_plot
```

### GCUH

```{r, fig.width=8, fig.height=7}
gcuh = filter(to_plot, hospital=='GCUH')
# redo plot for greater control
alt_plot = ggplot(gcuh, aes(x=time, y=est, ymin=lower, ymax=upper, col=group)) +
  geom_ribbon(alpha=0.2) +
  geom_line(size = 1.05) +  # make lines thicker
  scale_y_continuous(limits=c(0,0.5)) + # 
  coord_cartesian(xlim=c(0,pred_time))+ # limit based on censoring
  theme_bw() +
  scale_color_manual(NULL, values=cbPalette) +
  xlab('Days from ward admission') +
  ylab('Cumulative probability') +
  theme(legend.position = c(0.15, 0.8)) +
  facet_wrap(~team)
alt_plot
```

# Calendar time scale in the time-to-event analyses

Below we use calendar time for the survival analyses. This is useful for spotting what times there were relative increases or decreases in the 4, 5, 6 outcomes.

```{r make_times, include=FALSE}
### make the intervention times on a calendar time
# starting date/time (usual care)
ref_datetime = ISOdatetime(year=2020, month=5, day=25, hour=0, min=0, sec=1, tz='Australia/Brisbane')
ref_datetime = as.numeric(ref_datetime)
# intervention times
load('data/date_changes.RData') # from 0_date_changes.R
date_changes = mutate(date_changes, hosp_ordered = factor(hospital, levels=c('TPCH','RBWH','GCUH')))
lines = select(date_changes, hospital, hosp_ordered, date_intervention) %>%
  mutate(line = as.POSIXct(date_intervention, tz='Australia/Brisbane'),
         line = as.numeric(line),
         time = (line - ref_datetime)/(60*60*24)) # difference from starting date
# add all teams to lines
all_teams = select(baseline, hospital, team) %>% unique()
lines = left_join(lines, all_teams, by='hospital')
# subtract 28 days from intervention time to remove 4 week establishment phase prior to intervention
lines = mutate(lines, time = time - 28)
```

## a) Time to clinician-led care review discussion

```{r}
# put data on alternative time scale 
calendar_time_care_review = calendar_time(indata=for_model_care_review, changes=date_changes) # see 99_functions.R

diff_plots = NULL
for (hosp in c('TPCH','RBWH','GCUH')){
  
  # loop through teams
  teams = filter(for_model_care_review, hospital==hosp) %>% # get all teams
    pull(team) %>%
    unique()
  
  for (this_team in teams){
    # run Cox model
    cox_model_care_review_calendar = coxph(Surv(time = start, time2=end, event = event== 'Yes') ~ I(age/5) + pt_sex, data = filter(calendar_time_care_review, hospital == hosp, team==this_team))
    #
    base = basehaz(cox_model_care_review_calendar, centered=TRUE) %>%
      mutate(hospital = hosp)
    # make data for difference plot:
    diff_plot = base_haz_difference(inbase = base)  %>%
      mutate(team = this_team)
    diff_plots = bind_rows(diff_plots, diff_plot)
    
  }
}
```

### i) TPCH

```{r, fig.width=8}
# get ticks
ticks = x_ticks(Hospital='TPCH', 
                tick_dates = c('2020-05-25','2020-10-13','2020-12-01','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='TPCH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='TPCH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```

### ii) RBWH

```{r, fig.width=8, fig.height=8}
# get ticks
ticks = x_ticks(Hospital='RBWH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-15','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='RBWH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='RBWH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```

### iii) GCUH

```{r, fig.width=8, fig.height=6}
# get ticks
ticks = x_ticks(Hospital='GCUH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-01','2021-02-16','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='GCUH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='GCUH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```



## b) Time to care directives documents

```{r}
# put data on alternative time scale 
calendar_time_care_directive = calendar_time(indata=for_model_care_directive, changes=date_changes) # see 99_functions.R

diff_plots = NULL
for (hosp in c('TPCH','RBWH','GCUH')){
  
  # loop through teams
  teams = filter(calendar_time_care_directive, hospital==hosp) %>% # get all teams
    pull(team) %>%
    unique()
  
  for (this_team in teams){
    # run Cox model
    cox_model_care_directive_calendar = coxph(Surv(time = start, time2=end, event = event== 'Yes') ~ I(age/5) + pt_sex, data = filter(calendar_time_care_directive, hospital == hosp, team==this_team))
    #
    base = basehaz(cox_model_care_directive_calendar, centered=TRUE) %>%
      mutate(hospital = hosp)
    # make data for difference plot:
    diff_plot = base_haz_difference(inbase = base)  %>%
      mutate(team = this_team)
    diff_plots = bind_rows(diff_plots, diff_plot)
    
  }
}
```

### i) TPCH

```{r, fig.width=8}
# get ticks
ticks = x_ticks(Hospital='TPCH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-01','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='TPCH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='TPCH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```


### ii) RBWH

```{r, fig.width=8, fig.height=8}
# get ticks
ticks = x_ticks(Hospital='RBWH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-01','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='RBWH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='RBWH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```

### iii) GCUH

```{r, fig.width=8, fig.height=6}
# get ticks
ticks = x_ticks(Hospital='GCUH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-01','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='GCUH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='GCUH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```

## c) Time to palliative care referral

```{r}
# put data on alternative time scale 
calendar_time_palliative = calendar_time(indata=for_model_palliative, changes=date_changes) # see 99_functions.R

diff_plots = NULL
for (hosp in c('TPCH','RBWH','GCUH')){
  
  # loop through teams
  teams = filter(calendar_time_palliative, hospital==hosp) %>% # get all teams
    pull(team) %>%
    unique()
  
  for (this_team in teams){
    # run Cox model
    cox_model_palliative_calendar = coxph(Surv(time = start, time2=end, event = event== 'Yes') ~ I(age/5) + pt_sex, data = filter(calendar_time_palliative, hospital == hosp, team==this_team))
    #
    base = basehaz(cox_model_palliative_calendar, centered=TRUE) %>%
      mutate(hospital = hosp)
    # make data for difference plot:
    diff_plot = base_haz_difference(inbase = base)  %>%
      mutate(team = this_team)
    diff_plots = bind_rows(diff_plots, diff_plot)
    
  }
}
```

### i) TPCH

```{r, fig.width=8}
# get ticks
ticks = x_ticks(Hospital='TPCH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-01','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='TPCH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='TPCH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```


### ii) RBWH

```{r, fig.width=8, fig.height=8}
# get ticks
ticks = x_ticks(Hospital='RBWH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-01','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='RBWH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='RBWH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```

### iii) GCUH

```{r, fig.width=8, fig.height=6}
# get ticks
ticks = x_ticks(Hospital='GCUH', 
                tick_dates = c('2020-05-25','2020-09-01','2020-12-01','2021-03-01','2021-06-01'),
                date_changes = date_changes)
# plot
bplot = ggplot(data=filter(diff_plots, hospital=='GCUH'), aes(x=time, y=diff, col=team)) +
  geom_vline(data=filter(lines, hospital=='GCUH'), aes(xintercept=time), lty=2) + # vertical reference line at change
  geom_line() +
  geom_hline(yintercept=0, col='grey33', lty=3)+
  scale_color_manual(NULL, values=cbPalette) +
  scale_x_continuous(breaks=ticks$ticks, labels=ticks$labels)+
  xlab('Calendar time') +
  ylab('Cumulative hazard difference') +
  facet_wrap(~team, scales='free_y') +
  g.theme+
  theme(legend.position='none',
        axis.text.x = element_text(angle = 45, hjust=1))
bplot
```
