---
title: 'InterACT study: hospital-level summary for teams or hospitals'
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
  toc: false
reference_docx: rmarkdown-style-teams.docx
---
  
```{r setup, include=FALSE}
# this program is called by:
# 1_create_multiple_reports_teams.R

# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(dplyr.summarise.inform=FALSE) # turn off annoying warning
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
source('99_functions.R')
#library(pander)
library(dplyr)
library(stringr)
library(flextable) # for nice tables
library(visdat) # for missing data
library(summarytools)
# global options for summary tools
st_options(plain.ascii = FALSE,       # Always use this option in Rmd documents
           style = "rmarkdown",        # This too
           round.digits = 0, 
           headings = FALSE, 
           footnote = NA,             # Avoids footnotes which would clutter the results
           subtitle.emphasis = FALSE  # Improves layout with some rmarkdown themes
) 

# graphics things:
library(diagram)
library(ggplot2)
library(ggupset)  # to plot combination of symptoms
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7")

## for one hospital and team
names(baseline)[names(baseline) == team_var] = 'team' # rename
if(this_hospital != 'All'){ # do not filter if for all hospitals
  baseline = filter(baseline, 
                  hospital == this_hospital)
}
if(this_team != 'AllTeams'){
  baseline = filter(baseline, team %in% this_team)
}
# (other forms dealt with below)
```

Results for `r this_hospital` hospital, Team = `r this_team`. The data are from `r format(data_date, '%d-%b-%Y')`.

### Numbers screened by InterACT and at risk 

```{r, fig.width=3.8, fig.height=4.2}
N = nrow(baseline)  # starting number
# table of screening complete:
tab1 = table(baseline$baseline_screening_complete) 
ptab1 = round(prop.table(tab1)*100)
baseline = filter(baseline, baseline_screening_complete=='Yes')
# table of at risk conditional on screening complete:
tab2 = table(baseline$at_risk) 
ptab2 = round(prop.table(tab2)*100)
# table of 4,5,6 forms done
at_risk = filter(baseline, at_risk=='At risk')
tab3 = table(at_risk$complete_456) 
ptab3 = round(prop.table(tab3)*100)
# set up plotmat
M <- matrix(nrow = 7, ncol = 7, byrow = TRUE, data = 0)
# names
l1 = paste('Total\n', N, sep='')
l2 = paste('Screening\nnot complete\n', tab1[1], ' (', ptab1[1], '%)', sep='')
l3 = paste('Screening\ncomplete\n', tab1[2], ' (', ptab1[2], '%)', sep='')
l4 = paste('Not at risk\n', tab2[1], ' (', ptab2[1], '%)', sep='')
l5 = paste('At risk\n', tab2[2], ' (', ptab2[2], '%)', sep='')
l6 = paste('No follow-up forms\ncomplete\n', tab3[1], ' (', ptab3[1], '%)', sep='')
l7 = paste('At least one follow-\nup form complete\n', tab3[2], ' (', ptab3[2], '%)', sep='')
names = c(l1, l2, l3, l4, l5, l6, l7)
# transitions
M[2,1] = "' '"
M[3,1] = "' '"
M[4,3] = "' '"
M[5,3] = "' '"
M[6,5] = "' '"
M[7,5] = "' '"
colours = c("lightblue","green", "green", "yellow", "yellow",'pink','pink')
sizes = c(0.15, rep(0.2, 6)) # smaller first box
props = c(0.4, rep(0.48, 6)) # smaller first box
# output to markdown document
par(mai=rep(0.01,4))
plotmat(M, name = names, pos=c(1,2,2,2), curve = 0, box.cex = 1, box.lwd = 2,
        box.size = sizes, box.prop=props, arr.pos = 0.5, box.type = 'rect', arr.type = 'triangle', arr.length = 0.2, arr.width = 0.3, cex=0.25, # cex controls text size
        box.col = colours)
# export to figure
filename = paste('figures/', this_hospital, '/', str_remove_all(this_team, pattern='_| '), '/numbers.jpg', sep='')
jpeg(filename, width=4, height=4.8, quality = 100, res=500, units='in', bg='transparent')
par(mai=rep(0.01,4))
plotmat(M, name = names, pos=c(1,2,2,2), curve = 0, box.cex = 1, box.lwd = 2,
        box.size = sizes, box.prop=props, arr.pos = 0.5, box.type = 'rect', arr.type = 'triangle', arr.length = 0.2, arr.width = 0.3, cex=0.25, # cex controls text size
        box.col = colours)
invisible(dev.off())
```


```{r, include=FALSE}
### now just at risk patients ###
baseline = filter(baseline, 
                  at_risk == 'At risk', # must be at risk ...
                  baseline_screening_complete == 'Yes', # ... with baseline screening complete
                  #                  complete_456 == 'Yes', # at least one 4,5,6 form complete - do not condition on this
                  !is.na(at_risk))
# knock out patients not at risk:
ids = unique(baseline$participant_id)
clinicianled_review = filter(clinicianled_review, participant_id%in%ids) 
care_directive = filter(care_directive, participant_id%in%ids)
palliative_care_referral = filter(palliative_care_referral, participant_id%in%ids)

# work out numbers per week
n_total = nrow(baseline)
n_weeks = floor(as.numeric(max(baseline$admission_date, na.rm=TRUE) - min(baseline$admission_date, na.rm=TRUE))/7)
```

The `r n_total` patients at risk were collected over `r n_weeks` weeks, which is an average of `r round(n_total/n_weeks)` patients per week.

All results from now are for the `r nrow(baseline)` at risk.

### Presence of a documented care directive prior to this hospital admission

```{r, results='asis'}
# `change_var` is previously `care_review`; renamed because of function
with(baseline, freq(presence_eol_plan, round.digits=0, cumul = FALSE, report.nas =FALSE))
```

# SPICT 

### Ten most common combinations

```{r, fig.width=7, fig.height=3.25}
# get combinations of yes responses
yes_responses <-  dplyr::select(baseline, participant_id, spict.vars)  %>%
  as_tibble() %>%
  tidyr::gather(key='spict', value='response', -participant_id) %>%
  filter(response=='Yes') %>% # just yes
  select(-response) %>%
  mutate(spict = nice.rename.spict(spict)  )
# now make list from group responses
list_resp = group_by(yes_responses, participant_id) %>%
  summarise(spicts = list(spict))
# add zeros
all_ids = unique(baseline$participant_id)
missing.ids = all_ids[all_ids %in% list_resp$participant_id == FALSE]
zeros = as_tibble(data.frame(participant_id=missing.ids))
list_resp = bind_rows(list_resp, zeros)
# plot
cplot = ggplot(list_resp, aes(x = spicts)) +
  geom_bar(aes(y=..count..), fill = "indianred3") +
  g.theme +
  xlab("") +
  ylab("Number of patients") +
  scale_x_upset(n_intersections = 10, n_sets=10)  # top ten
#  theme(plot.margin = margin(t = 3, r = 1, b = 0, l = 90, unit = "pt"))
cplot
# export to figure
filename = paste('figures/', this_hospital, '/', str_remove_all(this_team, pattern='_| '), '/spict_upset.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(cplot)
invisible(dev.off())
```

### SPICT score

```{r, fig.width=7, fig.height=3.25}
to.bar = group_by(baseline, hospital, spict_score) %>%
  summarise(count = n()) %>%
  mutate(fill = as.factor(spict_score >= thresholds$spict_threshold))
spict.bar = ggplot(data=to.bar, aes(x=spict_score, y=count, fill=fill))+ # colour bars by positive
  xlab('SPICT score')+
  ylab('Number of patients')+
  scale_x_continuous(breaks=0:6)+
  geom_bar(stat='identity')+
  scale_fill_manual(NULL, values=c('pink','dark red'))+
  g.theme+
  theme(legend.position = 'none')
spict.bar
# export to figure
filename = paste('figures/', this_hospital, '/', str_remove_all(this_team, pattern='_| '), '/spict_bar.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(spict.bar)
invisible(dev.off())
```

### Numbers positive

The threshold for SPICT positive is `r thresholds$spict_threshold` or more.

```{r, results='asis'}
baseline = mutate(baseline, 
                  `SPICT positive` = as.numeric(spict_score >= thresholds$spict_threshold),
                  `SPICT positive` = factor(`SPICT positive`, levels=0:1, labels=c('No','Yes')))
with(baseline, freq(`SPICT positive`, cumul = FALSE, report.nas = FALSE))
```

\pagebreak

# CriSTAL 

### Ten most common combinations

```{r, fig.width=7, fig.height=3.25}
# get combinations of yes responses
yes_responses <-  dplyr::select(baseline, participant_id, cristal.vars)  %>%
  mutate(
    #age = ifelse(age>=75, 'Yes', 'No'), # convert to binary - no longer using age in cristal
    cristal_cfs_score1 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold, 'Yes', 'No'), # frailty score over 5
    cristal_cfs_score2 = ifelse(cristal_cfs_score >= thresholds$frailty_threshold2, 'Yes', 'No')) %>%  # frailty score over 7
  select(-cristal_cfs_score) %>% # no longer needed
  mutate_if(is.factor, as.character) %>% # convert all yes/no factors to characters
  as_tibble() %>%
  tidyr::gather(key='cristal', value='response', -participant_id) %>%
  filter(response=='Yes') %>% # just yes
  select(-response) %>%
  mutate(cristal = nice.rename.cristal(cristal)  )
# now make list from group responses
list_resp = group_by(yes_responses, participant_id) %>%
  summarise(cristals = list(cristal))
# add zeros
all_ids = unique(baseline$participant_id)
missing.ids = all_ids[all_ids %in% list_resp$participant_id == FALSE]
zeros = as_tibble(data.frame(participant_id=missing.ids))
list_resp = bind_rows(list_resp, zeros)
# plot
cplot = ggplot(list_resp, aes(x = cristals)) +
  geom_bar(aes(y=..count..), fill = "blue2") +
  g.theme + 
  xlab("") +
  ylab("Number of patients") +
  scale_x_upset(n_intersections = 10, n_sets=10) # top ten
cplot
filename = paste('figures/', this_hospital, '/', str_remove_all(this_team, pattern='_| '), '/cristal_upset.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(cplot)
invisible(dev.off())
```

### CriSTAL score 

```{r, fig.width=7, fig.height=3.25}
to.bar = group_by(baseline, hospital, cristal_score) %>%
  summarise(count = n()) %>%
  mutate(fill = as.factor(cristal_score >= thresholds$cristal_threshold))
cristal.bar = ggplot(data=to.bar, aes(x=cristal_score, y=count, fill=fill))+ # colour bars by positive
  xlab('CriSTAL score')+
  ylab('Number of patients')+
  scale_x_continuous(breaks=0:10)+
  #  scale_y_continuous()
  geom_bar(stat='identity')+
  scale_fill_manual(NULL, values=c('skyblue','dark blue'))+
  g.theme+
  theme(legend.position = 'none')
cristal.bar
# export to figure
filename = paste('figures/', this_hospital, '/', str_remove_all(this_team, pattern='_| '), '/cristal_bar.jpg', sep='')
jpeg(filename, width=5.5, height=4, quality = 100, res=500, units='in')
print(cristal.bar)
invisible(dev.off())
```

### Numbers positive

A CriSTAL positive score is `r thresholds$cristal_threshold` or above.

```{r, results='asis'}
baseline = mutate(baseline, 
                  `CriSTAL positive` = as.numeric(cristal_score >= thresholds$cristal_threshold),
                  `CriSTAL positive` = factor(`CriSTAL positive`, levels=0:1, labels=c('No','Yes')))
with(baseline, freq(`CriSTAL positive`, cumul=FALSE, report.nas = FALSE))
```


\pagebreak

# Outcomes

The results below exclude those with incomplete follow-up forms.

```{r, results = 'asis'}
outcome_data = bind_rows(clinicianled_review,
                      care_directive,
                      palliative_care_referral, .id='outcome') %>%
  left_join(select(baseline, participant_id, spict_unplan_hosp), by='participant_id') %>% # add back unplanned adm
  select(outcome, change_var, spict_unplan_hosp) %>%
  filter(!is.na(change_var)) %>% # remove few missing
  mutate(outcome = factor(as.numeric(outcome), levels=1:3,
                          labels=c('Clinician-led care review discussion','Review of any care directives','Referred to palliative care')))

## now make table
one_table = group_by(outcome_data, outcome, change_var) %>%
  summarise(n = n()) %>%
  group_by(outcome) %>%
  mutate(pct = round(100*prop.table(n)),
         pct = paste(pct, '%', sep=''),
         cell = paste(n, ' (', pct, ')', sep='')) %>%
  select(outcome, change_var, cell) %>%
  group_by(outcome) %>%
  tidyr::spread(change_var, cell) %>%
  mutate(Prior = ifelse(is.na(Prior), '0 (0%)', Prior),
         Yes = ifelse(is.na(Yes), '0 (0%)', Yes))
# total column
total_column = group_by(outcome_data, outcome) %>%
  summarise(Total = n()) %>%
  mutate(Total = paste(Total, '(100%)')) %>%
  ungroup()
# combine table and totals
one_table = left_join(one_table, total_column, by='outcome')
# make table
ft = flextable(one_table) 
autofit(ft)
```

"Prior" is prior to this admission.

### Bar-plot of the three outcomes

```{r, fig.width=7, fig.height=4}
#
stats = group_by(outcome_data, outcome, change_var) %>%
  summarise(n = n()) %>%
  group_by(outcome) %>%
  mutate(pct = round(100*prop.table(n)),
         pct = paste(pct, '%'))
# plot
cplot = ggplot(data=outcome_data, aes(x=change_var, fill=change_var))+
  geom_bar()+
  scale_fill_manual(NULL, values=c("#66CDAA", "#1C86EE", "#C1CDCD"))+
  xlab('')+
  ylab('Number of patients')+
  g.theme +
  geom_text(data=stats, aes(x=change_var, y=n, label=pct), col=grey(0.1), nudge_y=1)+ # nudge to push text onto bars
  facet_wrap(~outcome)+
  theme(legend.position = 'none')
cplot
filename = paste('figures/', this_hospital, '/', str_remove_all(this_team, pattern='_| '), '/outcomes_bar.jpg', sep='')
jpeg(filename, width=6.7, height=4, quality = 100, res=500, units='in', bg='transparent')
print(cplot)
invisible(dev.off())
```


